Lesson 5
========================================================

### Multivariate Data
Notes:

***

### Moira Perceived Audience Size Colored by Age
Notes:

***

### Third Qualitative Variable
Notes:

```{r Third Qualitative Variable}
ggplot(aes(x = gender, y = age),
       data = subset(pf, !is.na(gender))) + geom_boxplot()
```

***

### Plotting Conditional Summaries
Notes:

```{r Plotting Conditional Summaries}
#age_gender_groups <- group_by(pf, age, gender)

#pf.fc_by_age_gender <- summarise(age_gender_groups,
 #                                mean_friend_count = mean(friend_count),
  #                               median_friend_count = median(friend_count),
   #                              n = n())
#pf.fc_by_age_gender <- ungroup(pf.fc_by_age_gender)
#pf.fc_by_age_gender <- arrange(pf.fc_by_age_gender, age)
#pf.fc_by_age_gender <- filter(pf.fc_by_age_gender, !is.na(gender))



pf.fc_by_age_gender2 <- pf %>%
  filter(!is.na(gender)) %>%
  group_by(age, gender) %>%
  summarise(mean_friend_count = mean(friend_count),
            median_friend_count = median(friend_count),
            n = n()) %>%
  ungroup() %>%
  arrange(age)
head(pf.fc_by_age_gender2)
tail(pf.fc_by_age_gender2)

ggplot(aes(age, median_friend_count), data = pf.fc_by_age_gender) + geom_line(aes(color = gender))
```

***

### Thinking in Ratios
Notes:

***

### Wide and Long Format
Notes:

***

### Reshaping Data
Notes:

```{r}
install.packages('reshape2')
library(reshape2)
pf.fc_by_age_gender.wide = dcast(pf.fc_by_age_gender,
                                 age ~ gender,
                                 value.var = "median_friend_count")
head(pf.fc_by_age_gender.wide)
```


***

### Ratio Plot
Notes:

```{r Ratio Plot}
ggplot(aes(age, female/male), data = pf.fc_by_age_gender.wide) + geom_line() +
  geom_hline(yintercept = 1, linetype = 2)
```

***

### Third Quantitative Variable
Notes:

```{r Third Quantitative Variable}
pf$year_joined <- floor(2014 - (pf$tenure/365))
```

***

### Cut a Variable
Notes:

```{r Cut a Variable}
?cut
pf$year_joined.buckets <- cut(pf$year_joined, c(2004, 2009, 2011, 2012, 2014))

```

***

### Plotting it All Together
Notes:

```{r Plotting it All Together}
ggplot(aes(x = age, y = friend_count),
       data = subset(pf, !is.na(year_joined.buckets))) +
  geom_line(aes(color = year_joined.buckets), stat = 'summary', fun.y = median)
```

***

### Plot the Grand Mean
Notes:

```{r Plot the Grand Mean}
ggplot(aes(x = age, y = friend_count),
       data = subset(pf, !is.na(year_joined.buckets))) +
  geom_line(aes(color = year_joined.buckets), stat = 'summary', fun.y = mean) +
  geom_line(stat = 'summary', fun.y = mean, linetype = 6)
```

***

### Friending Rate
Notes:

```{r Friending Rate}
with(subset(pf, tenure >= 1), summary(friend_count / tenure))
```

***

### Friendships Initiated
Notes:

What is the median friend rate?

What is the maximum friend rate?

```{r Friendships Initiated}

```

***

### Bias-Variance Tradeoff Revisited
Notes:

```{r Bias-Variance Tradeoff Revisited}

ggplot(aes(x = tenure, y = friendships_initiated / tenure),
       data = subset(pf, tenure >= 1)) +
  geom_line(aes(color = year_joined.buckets),
            stat = 'summary',
            fun.y = mean)

ggplot(aes(x = 7 * round(tenure / 7), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

ggplot(aes(x = 30 * round(tenure / 30), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

ggplot(aes(x = 90 * round(tenure / 90), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

ggplot(aes(x = tenure, y = friendships_initiated / tenure),
       data = subset(pf, tenure >= 1)) +
  geom_smooth(aes(color = year_joined.buckets))

```

***

### Sean's NFL Fan Sentiment Study
Notes:

***

### Introducing the Yogurt Data Set
Notes:

***

### Histograms Revisited
Notes:

```{r Histograms Revisited}
getwd()
yo <- read.csv("yogurt.csv")
str(yo)

yo$id <- factor(yo$id)
str(yo)

library(ggplot2)
ggplot(aes(price), data = yo) + geom_histogram(binwidth = 1)
```

***

### Number of Purchases
Notes:

```{r Number of Purchases}
?transform
names(yo)
yo <- transform(yo, all.purchases = strawberry + blueberry + pina.colada + plain + mixed.berry)
summary(yo$all.purchases)
```

***

### Prices over Time
Notes:

```{r Prices over Time}
summary(yo)
ggplot(aes(time, price), data = yo) + geom_point(alpha = .1)
```

***

### Sampling Observations
Notes:

***

### Looking at Samples of Households

```{r Looking at Sample of Households}
sample.ids <- sample(levels(yo$id), 16)
sample.ids
sample.ids2 <- sample(levels(yo$id), 9)

ggplot(aes(time, price), data = subset(yo, id %in% sample.ids2)) + 
  geom_point(aes(size = all.purchases), pch = 1) +
    facet_wrap( ~ id) +
    geom_line() 
```

***

### The Limits of Cross Sectional Data
Notes:

***

### Many Variables
Notes:

***

### Scatterplot Matrix
Notes:

***

### Even More Variables
Notes:

***

### Heat Maps
Notes:

```{r}
nci <- read.table("nci.tsv")
colnames(nci) <- c(1:64)
```

```{r}
nci.long.samp <- melt(as.matrix(nci[1:200,]))
names(nci.long.samp) <- c("gene", "case", "value")
head(nci.long.samp)

ggplot(aes(y = gene, x = case, fill = value),
  data = nci.long.samp) +
  geom_tile() +
  scale_fill_gradientn(colours = colorRampPalette(c("blue", "red"))(100))
```


***

### Analyzing Three of More Variables
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

